---
title: "Tidy data up"
author: "Luis Fernando Delgado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    theme: paper
    df_print: tibble
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
fig.dim = c(7, 4)
library(ggrepel)
```

# Load data
```{r}
# Sourcing External R Script
source("https://raw.githubusercontent.com/Cassava2050/PPD/main/utilities_tidy.R")

# Setting Up File Paths and Parameters
folder <- here::here("data//")  
file <- "phenotype.csv"
skip_col <- 3 
trial_interest <- "MDEPR"
year_interest <- 2022

# Loading Data
# The function 'read_cassavabase' should be defined in the sourced script
sel_data <- read_cassavabase(phenotypeFile = paste0(folder, file))
```

# Standardizing Column Names
The function 'change_colname' should also be defined in the sourced script

```{r}
sel_data_kp <- change_colname(sel_data, NA)
```
# Changing Column type

```{r}
obs_col <- c(
  names(sel_data_kp)[str_detect(names(sel_data_kp), "obs_")],
  "use_rep_number", "blockNumber",
  "use_plot_number", "use_plot_width",
  "use_plot_length"
)
```

# mutate all traits into numeric

```{r}
sel_data_kp <- sel_data_kp %>% 
  mutate(across(all_of(obs_col), as.numeric))


# remove - , replace by _
names(sel_data_kp) = gsub("-", "_", names(sel_data_kp))

# remove trials: 202301, 202304, 202307. These were analyzed 
sel_data_kp <- 
  sel_data_kp %>% filter(!str_detect(use_trial_name, "202301|202304|202307"))


# Duplications in row and cols
duplicated_plot <- row_col_dup(sel_data_kp)
```
# Plot trial layout
```{r}
trial_layout(sel_data_kp)
```

# Check the clone name

```{r}
cloneName_new_old <- check_clone_name(
  clone_list = sel_data_kp$use_accession_name,
  new_names = NA,
  add_check = NULL
)

trial_standard <- sel_data_kp %>%
  left_join(cloneName_new_old,
            by = c("use_accession_name" = "accession_name_ori")
  ) %>%
  select(-use_accession_name) %>%
  rename(use_accession_name = use_accession_name.y)
```
# Add GIS data

```{r}
trial_standard <- add_GIS(trial_standard)

# Get world map data for reference
vietnam_map <- map_data("world", region = "Vietnam")

# Create a data frame with your locations
locations <- data.frame(
  Location = c("Tay Ninh", "Dong Nai"),
  Latitude = c(11.3009, 11.1432),
  Longitude = c(106.1107, 107.2742)
)

# Create a ggplot object for the map
static_map <- ggplot() +
  geom_polygon(data = vietnam_map, aes(x = long, y = lat, group = group), 
               fill = "lightgray", col = "black", linewidth = 0.03) +
  geom_point(data = locations, aes(x = Longitude, y = Latitude), size = 2, color = "red") +
  geom_text_repel(data = locations, aes(x = Longitude, y = Latitude, label = Location), 
                  size = 3) +
  geom_text(aes(x = 108.5, y = 10, label = "Author: Luis Fdo. Delgado"), size = 1.5) +
  labs(x =  "Latitude", y = "Longitude") +
  labs(title = "Locations in Vietnam", subtitle = "Tay Ninh and Dong Nai") +
  theme(plot.title = element_text(face = "bold.italic"),
        plot.subtitle = element_text(face = "italic")) +
  coord_fixed(1.3) # This sets the aspect ratio

# Display the map
print(static_map)

ggsave(paste("images\\map", trial_interest, ".png", sep = "_"),
       plot = static_map, units = "in", dpi = 300, width = 6, height = 5
)

```
# Extracting checks

```{r max.print=1}
accession_rep_ct <- trial_standard %>%
  count(use_trial_name, use_accession_name, use_rep_number)  %>%
  arrange(use_trial_name) %>%
  filter(n>1)
print(accession_rep_ct) 
```

# Harvesting time

```{r max.print=1}
conducted_trials <- 
  trial_standard %>% group_by(use_trial_name, use_plant_date,use_harvest_date, use_location) %>% 
  summarise(n_gen = n_distinct(use_accession_name)) %>% 
  mutate(harvesting_time = 
           interval(ymd(use_plant_date), ymd(use_harvest_date)) %>% as.period,
         harvesting_time = paste0(harvesting_time@month, "month ", harvesting_time@day, "day")) %>% 
  ungroup()

print(conducted_trials %>% rmarkdown::paged_table())
```
# plot plant number

```{r max.print=1}
plants_plot <- trial_standard %>%
  group_by(use_trial_name) %>%
  count(obs_planted_number_plot) 
print(plants_plot)
```

# Frequency harvest plant number

```{r}
plants_harvested <- trial_standard %>%
  group_by(use_trial_name) %>%
  count(obs_harvest_number) %>% arrange(desc(obs_harvest_number))

# planted and harvested
plants_plot %>% select(-n) %>% 
  left_join(plants_harvested %>% 
              summarise(harvested_plants = max(obs_harvest_number)), by = "use_trial_name") %>% 
  write.table("clipboard", sep="\t", col.names = T, row.names = F)
```

# Double check yield calculation

```{r fig.height=6, fig.width=8}
plants_to_harvest <- plants_harvested %>% 
  ggplot(aes(x = factor(obs_harvest_number), 
             y = n, fill = factor(obs_harvest_number))) +
  geom_col(col = 'black') +
  #scale_fill_jco() +
  #theme_xiaofei() +
  theme(axis.text.x = element_text(vjust = 1, angle = 65),
        text = element_text(size = 20),
        legend.position="top")+
  labs(x = "Harvest_plant_number", y = "Freq", fill = "Harvest_plant_number") +
  facet_wrap(~ use_trial_name)

print(plants_to_harvest)

ggsave(paste("images\\bar", trial_interest, ".png", sep = "_"),
       plot = plants_to_harvest, units = "in", dpi = 300, width = 9, height = 6)
```

# Compute germination, yield, yield_starch

```{r}
trial_standard <- trial_standard %>%
  mutate(obs_harvest_number_plan =
           case_when(str_detect(use_trial_name, "2023111")  ~ 4,
                     str_detect(use_trial_name, "2023112")  ~ 4,
                     str_detect(use_trial_name, "2023113")  ~ 10,
                     str_detect(use_trial_name, "2023114")  ~ 10),
         obs_germination_perc = obs_germinated_number_plot/obs_planted_number_plot * 100,
         # 2) calculate area per plant
         area_plant = (use_plot_length*use_plot_width)/obs_planted_number_plot,
         # 3) calculate the yield_v4 obs_planted_number_plot
         obs_yield_ha = ifelse(obs_yield_ha == 0, NA, obs_yield_ha),
         obs_starch_content = ifelse(obs_starch_content == 0, NA, obs_starch_content),
         obs_yield_ha_v2 = (((obs_root_weight_plot*10000)/(area_plant*obs_harvest_number_plan))/1000),
         obs_starch_yield_ha = obs_starch_content * obs_yield_ha_v2 / 100)

library(plotly)

p1 <- trial_standard %>% ggplot() +
  geom_point(aes(x = obs_yield_ha, y = obs_yield_ha_v2, color = use_plot_number), show.legend = F) +
  facet_wrap(~use_trial_name) +
  theme_xiaofei()

ggplotly(p1)
detach("package:plotly", unload = TRUE)
```











